---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcaiman.registry

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`rcaiman.registry` provides the infrastructure used by `rcaiman` to
describe imaging systems used in canopy photography.

The package implements a structured registry where instrument
characteristics and interpretation rules are declared through
specifications combined into registry entries. 

The package also provides tools to export and read **registry snapshots**
(YAML files) that capture a portable subset of a registry entry.

## Installation

You can install the development version of rcaiman.registry like so:

``` r
# install.packages("devtools")
devtools::install_github("GastonMauroDiaz/rcaiman.registry")
```

## Registry entries

A registry entry represents a complete technical description of an imaging
system. It typically combines:

- an `embedded_metadata_sig` describing how metadata is detected in files
- a `geometry_spec` defining the directional reference frame of the image
- a `radiometry_spec` defining how pixel values should be interpreted


## How to create and navigate a registry entry

The package provides helper functions that allow you to build a valid
registry entry step by step:

```{r include=FALSE}
library(rcaiman.registry)
```


```{r example}
foo <- new_registry_entry(
  "Nikon_Coolpix5700.FCE9.CIEFAP",
  body = "E5700",
  body_manufacturer = "NIKON CORP",
  body_serial = "7053067",
  lens = "Zoom Nikkor ED 8.9-71.2mm 1:2.8-4.2",
  lens_manufacturer = "NIKON CORP",
  auxiliary_lens = "Fisheye Converter FC-E9 0.2x",
  auxiliary_lens_manufacturer = "NIKON CORP",
  institution = "CIEFAP"
)

foo <- add_file_sig(
  foo,
  id = "raw",
  extension = "NEF",
  filename_pattern = "^DSCN[0-9]{4}$"
)

foo <- add_embedded_metadata_sig(
  foo,
  id = "exif_01",
  namespace = "exif",
  dim = c(1288, 962),
  rules = list(
    "Camera Model Name" = "E5700",
    "Software" = "E5700v1.1",
    "CFA Pattern" = "[Yellow,Cyan][Green,Magenta]",
    "Compression" = "Uncompressed",
    "Bits Per Sample" = "12",
    "Quality" = "RAW"
  ),
  file_sig = "raw"
)

foo <- add_geometry_spec(
  foo,
  embedded_metadata_sig = "exif_01",
  id = "simple_method",
  model = "radial_projection",
  parameters = c(0.6380,  0.0307, -0.0200),
  zenith_col_row = c(645, 494),
  horizon_radius = 378,
  is_horizon_circle_clipped = FALSE,
  max_zenith_angle = 94.8,
  notes = "Calibration documented in doi:10.1016/j.agrformet.2024.110020",
  contact_information = "gastonmaurodiaz@gmail.com"
)

foo <- add_radiometry_spec(
  foo,
  embedded_metadata_sig = "exif_01",
  geometry_spec = "simple_method",
  id = "spectral_bands",
  type = "interpretive_constraint",
  cfa_pattern = matrix(c("Green", "Yellow",
                         "Magenta", "Cyan"), byrow = TRUE, ncol = 2),
  spectral_mapping = list(Red = "(Yellow + Magenta)/2",
                          Green = "Green",
                          Blue = "Cyan"),
  offset_value = list("100" = 0),
  contact_information = "gastonmaurodiaz@gmail.com"
)
```

A registry entry available in memory can be quickly inspected to
understand which instrument it describes:

```{r print}
foo
```
`summary()` can be used to list the specification identifiers.

```{r summary}
summary(foo)
```
The `get_*()` helpers allow specifications to be retrieved easily for
use in scripts that interact with `rcaiman`.

```{r}
fs <- get_file_sig(foo, "raw")
em <- get_embedded_metadata_sig(foo, "exif_01")
fs <- get_file_sig(foo, em$file_sig)
spec <- get_geometry_spec(foo, "exif_01", "simple_method")
spec <- get_geometry_spec(em, "simple_method")
spec <- get_geometry_spec(foo, "exif_01", "simple_method")
rspec <- get_radiometry_spec(spec, "spectral_bands")
rspec <- get_radiometry_spec(foo, "exif_01", "simple_method", "spectral_bands")
```


A registry snapshot can be exported as a YAML file describing the
geometric projection and radiometric interpretation of the imaging
system:

```r
export_registry_snapshot(
  registry_entry = foo,
  embedded_metadata_sig = "exif_01",
  geometry_spec = "simple_method",
  radiometry_spec = "spectral_bands",
  file = "registry_snapshot.yaml"
)
```

## On calibration

The functions `calibrate_lens()` and `extract_radiometry()` provide
support for implementing the simple calibration method described in:

Díaz, G. M., Lang, M., & Kaha, M. (2024).  
*Simple calibration of fisheye lenses for hemispherical photography of the forest canopy.*  
Agricultural and Forest Meteorology, 352, 110020.  
https://doi.org/10.1016/j.agrformet.2024.110020


## Registry architecture

The registry framework can be organised as a small modular suite of
packages:

rcaiman.registry          → registry framework  
rcaiman.registry.data     → registry entries maintained with the project  
rcaiman.registry.<lab>    → registries maintained by individual groups

This arrangement allows different research groups to maintain and share
their own imaging system definitions while remaining compatible with
the broader `rcaiman` ecosystem.

## Extending the registry ecosystem

The registry framework is intentionally flexible. In addition to the
registry entries distributed in rcaiman.registry.data (currently under
development), research groups may choose to distribute their own
registry definitions through dedicated R packages.

One simple naming convention that can help keep things organised is:

`rcaiman.registry.<lab_or_project>`

Packages following this pattern would typically depend on
`rcaiman.registry` and contain custom `hs_registry_entry` definitions or
YAML snapshots describing specific imaging systems.
