---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rcaiman.registry

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`rcaiman.registry` provides the infrastructure used by `rcaiman` to
describe imaging systems used in canopy photography.

The package implements a structured registry where instrument
characteristics and interpretation rules are declared through
specifications combined into registry entries. 

The package also provides tools to export and read **registry snapshots**
(YAML files) that capture a portable subset of a registry entry.

## Installation

You can install the development version of rcaiman.registry like so:

``` r
# install.packages("devtools")
devtools::install_github("GastonMauroDiaz/rcaiman.registry")
```

## Registry entries

A registry entry represents a complete technical description of an imaging
system. It typically combines:

- an `embedded_metadata_sig` describing how metadata is detected in files
- a `geometry_spec` defining the directional reference frame of the image
- a `radiometry_spec` defining how pixel values should be interpreted


## How to create and navigate a registry entry

The package provides helper functions that allow you to build a valid
registry entry step by step:

```{r include=FALSE}
library(rcaiman.registry)
```


```{r example}
foo <- new_registry_entry(
  "Nikon_Coolpix5700.FCE9.CIEFAP",
  body = "E5700",
  body_manufacturer = "NIKON CORP",
  body_serial = "7053067",
  lens = "Zoom Nikkor ED 8.9-71.2mm 1:2.8-4.2",
  lens_manufacturer = "NIKON CORP",
  auxiliary_lens = "Fisheye Converter FC-E9 0.2x",
  auxiliary_lens_manufacturer = "NIKON CORP",
  institution = "CIEFAP"
)

foo <- add_file_sig(
  foo,
  id = "raw",
  extension = "NEF",
  filename_pattern = "^DSCN[0-9]{4}$"
)

foo <- add_embedded_metadata_sig(
  foo,
  id = "exif_01",
  namespace = "exif",
  dim = c(1288, 962),
  validated_with_rawpy = "0.19.0",
  validated_with_libraw = "0.21.1",
  rules = list(
    "Camera Model Name" = "E5700",
    "Software" = "E5700v1.1",
    "CFA Pattern" = "[Yellow,Cyan][Green,Magenta]",
    "Compression" = "Uncompressed",
    "Bits Per Sample" = "12",
    "Quality" = "RAW"
  ),
  file_sig = "raw"
)

foo <- add_geometry_spec(
  foo,
  embedded_metadata_sig = "exif_01",
  id = "simple_method",
  model = "radial_projection",
  parameters = c(0.6380,  0.0307, -0.0200),
  zenith_col_row = c(645, 494),
  horizon_radius = 378,
  is_horizon_circle_clipped = FALSE,
  max_zenith_angle = 94.8,
  notes = "Calibration documented in doi:10.1016/j.agrformet.2024.110020",
  contact_information = "gastonmaurodiaz@gmail.com"
)

foo <- add_radiometry_spec(
  foo,
  embedded_metadata_sig = "exif_01",
  geometry_spec = "simple_method",
  id = "spectral_bands",
  type = "interpretive_constraint",
  cfa_pattern = matrix(c("Green", "Yellow",
                         "Magenta", "Cyan"), byrow = TRUE, ncol = 2),
  spectral_mapping = list(Red = "(Yellow + Magenta)/2",
                          Green = "Green",
                          Blue = "Cyan"),
  offset_value = list("100" = 0),
  contact_information = "gastonmaurodiaz@gmail.com"
)

foo <- add_radiometry_spec(
  foo,
  embedded_metadata_sig = "exif_01",
  geometry_spec = "simple_method",
  id = "simple_method",
  type = "flat_field_correction",
  model = "flat_field_simple_polynomial",
  parameters = list("5.0" = c(0.0638, -0.101)),
  notes = "Calibration documented in doi:10.1016/j.agrformet.2024.110020",
  contact_information = "gastonmaurodiaz@gmail.com"
)
```

A registry entry available in memory can be quickly inspected to
understand which instrument it describes:

```{r print}
foo
```
`summary()` can be used to list the specification identifiers.

```{r summary}
summary(foo)
```
The `get_*()` helpers allow specifications to be retrieved easily for
use in scripts that interact with `rcaiman`.

```{r}
fs <- get_file_sig(foo, "raw")
ems <- get_embedded_metadata_sig(foo, "exif_01")
fs <- get_file_sig(foo, ems$file_sig)
gspec <- get_geometry_spec(foo, "exif_01", "simple_method")
gspec <- get_geometry_spec(ems, "simple_method")
rspec <- get_radiometry_spec(gspec, "spectral_bands")
rspec <- get_radiometry_spec(foo, "exif_01", "simple_method", "spectral_bands")
```


The registry describes imaging systems, while registry snapshots capture
the subset of metadata required to interpret individual processed images.

```r
export_registry_snapshot(
  registry_entry = foo,
  embedded_metadata_sig = "exif_01",
  geometry_spec = "simple_method",
  interpretive_constraint = "spectral_bands",
  flat_field_correction = simple_method,
  file = "registry_snapshot.yaml"
)
```

## On calibration

The functions `calibrate_lens()` and `extract_radiometry()` provide
support for implementing the simple calibration method described in:

Díaz, G. M., Lang, M., & Kaha, M. (2024).  
*Simple calibration of fisheye lenses for hemispherical photography of the forest canopy.*  
Agricultural and Forest Meteorology, 352, 110020.  
https://doi.org/10.1016/j.agrformet.2024.110020


## Registry architecture

The registry project can be organised as a small modular suite of
packages:

rcaiman.registry          → registry framework

rcaiman.registry.data     → registry entries maintained by \pkg{rcaiman} maintainer

rcaiman.registry.<foo>    → registry entries maintained by individual groups

## Extending the registry

The `rcaiman.registry.<foo>` naming convention improves discoverability
and signals that the package extends the registry project. Replacing
`<foo>` with a lab or project name provides an intuitive naming scheme.

Packages following this pattern would typically depend on
`rcaiman.registry` and contain `hs_registry_entry` objects
describing specific imaging systems. Alternatively, when resources for
creating a full R package are not available, a simple repository
containing YAML snapshot files can also be used.

These extension packages are encouraged to distribute:

- the registry entry describing the imaging system and its specifications 
- raw image data
- the corresponding preprocessed output

Providing pairs of raw and preprocessed data enables testing,
validation, and reproducibility of registry entries.

GitHub repositories provide a convenient distribution channel for these
packages, particularly given the size limitations imposed by CRAN for
packages containing example datasets.
